---
title: "DATA 624 - HOMEWORK 7"
author: "Fan Xu"
date: "4/18/2021"
output:
  rmdformats::readthedown:
    code_folding: hide
    df_print: paged
    highlight: tango
    number_sections: no
    smooth_scroll: yes
    theme: united
    toc_collapsed: yes
    toc_depth: 5
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
  html_notebook: default
  pdf_document:
    extra_dependencies:
    - geometry
    - multicol
    - multirow
  word_document:
    toc: yes
    toc_depth: '5'
theme: lumen
number_sections: yes
toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load package, message=FALSE, warning=FALSE}
library(tidyverse)
library(caret)
library(pls)
library(missForest)
library(elasticnet)
library(corrplot)
```



# Question 6.2 

**Developing a model to predict permeability (see Sect. 1.4) could save significant resources for a pharmaceutical company, while at the same time more rapidly identifying molecules that have a sufficient permeability to become a drug:**

## (a)
Start R and use these commands to load the data:

```{r Q6.2 load data}
library(AppliedPredictiveModeling)
data(permeability)
```

The matrix `fingerprints` contains the 1,107 binary molecular predictors for the 165 compounds, while `permeability` contains permeability response.


## (b) 
The fingerprint predictors indicate the presence or absence of substructures of a molecule and are often sparse meaning that relatively few of the molecules contain each substructure. Filter out the predictors that have low frequencies using the `nearZeroVar` function from the caret package. How many predictors are left for modeling?

**Answer:**
388 predictors are left for modeling after removing predictors with near zero variance.

```{r near zeo variance filter}
col_nearzero <- nearZeroVar(fingerprints)

fp_filtered <- fingerprints[,-col_nearzero]

ncol(fp_filtered)
```

## (c)
Split the data into a training and a test set, pre-process the data, and tune a PLS model. How many latent variables are optimal and what is the corresponding resampled estimate of R2?



### Train-test-split
```{r Q6.2 train test split}
set.seed(0)

train_index <- createDataPartition(permeability,
                                   p = 0.8,
                                   list = FALSE,
                                   times = 1) %>%
  as.vector()

data_train_X <- fp_filtered[train_index,] 
data_train_Y <- permeability[train_index,]
data_test_X <- fp_filtered[-train_index,] 
data_test_Y <- permeability[-train_index,]
```



### Tune PLS model

**Answer:**

1. 10 latent variables are optimal;

2. The corresplonding resampled estimate of R2 is 0.8433209.
```{r Q6.2 tune pls model}
set.seed(0)

ctrl <- trainControl(method = "cv", number = 10)


plsTune <- train(data_train_X, data_train_Y,
                 method = 'pls',
                 tuneLength = 20,
                 trControl = ctrl,
                 preProc = c('center', 'scale'))
resampled_pred <- predict(plsTune, data_train_X)

plsTune$bestTune  

postResample(resampled_pred, data_train_Y)
```

### Variance Important Evaluation

```{r Q6.2 variance important evaluation}
plot(varImp(plsTune), top = 20, main = 'Variance Important Score (Top 20)', ylab = 'Predictor')
```

## (d)

**Predict the response for the test set. What is the test set estimate of R2?**

The R2 of the test set is 0.3160511.
```{r Q6.2 predict response}
data_pred <- predict(plsTune, data_test_X)
pls_pred_metrics <- postResample(data_pred, data_test_Y)
pls_pred_metrics
```



## (e) 
**Try building other models discussed in this chapter. Do any have better predictive performance?**

**Answer:**
Ridge, Lasso and Elastic Net models are built as below. In summary, Elastic Net model has the best predictive performance among all.


### Ridge
The optimal lambda is 0.1473684, the R2 of test set is 0.3209984

#### Train Ridge Model
```{r Q6.2 ridge regression}
set.seed(0)
ridgeGrid <- data.frame(.lambda = seq(0, .2, length = 20))
RidgeTune <- train(data_train_X, data_train_Y,
                   method = 'ridge',
                   tuneGrid = ridgeGrid,
                   trControl = ctrl,
                   preProc = c('center', 'scale'))

RidgeTune
```

#### Prediction
```{r Q6.2 ridge prediction}
ridge_pred <- predict(RidgeTune, data_test_X)
ridge_pred_metrics <- postResample(ridge_pred, data_test_Y)
ridge_pred_metrics
```


### Lasso

#### Train Lasso Model
The Optimal fraction is 0.03368421,  the R2 of test set is 0.3603223
```{r Q6.2 lasso}
set.seed(0)
lassoGrid <- data.frame(.fraction = seq(0.01, .1, length = 20))
lassoTune <- train(data_train_X, data_train_Y,
                   method = 'lasso',
                   tuneGrid = lassoGrid,
                   trControl = ctrl,
                   preProc = c('center', 'scale'))

lassoTune


```




#### Prediction
```{r Q6.2 lasso prediction}
lasso_pred <- predict(lassoTune, data_test_X)
lasso_pred_metrics <- postResample(lasso_pred, data_test_Y)
lasso_pred_metrics
```


### Elastic Net

#### Train Elastic Net Model
The optimal fraction = 0.09052632 and lambda = 0.1789474, the R2 of test set is 0.3978472
```{r Q6.2 elastic net}
set.seed(0)
enetGrid <- data.frame(.lambda = seq(0, .2, length = 20),
                       .fraction = seq(0.01, .1, length = 20))
enetTune <- train(data_train_X, data_train_Y,
                   method = 'enet',
                   tuneGrid = enetGrid,
                   trControl = ctrl,
                   preProc = c('center', 'scale'))

enetTune
```


#### Prediction
```{r Q 6.2 elastic net predction}
enet_pred <- predict(enetTune, data_test_X)
enet_pred_metrics <- postResample(enet_pred, data_test_Y)
enet_pred_metrics

```

## (f) 
Would you recommend any of your models to replace the permeability laboratory experiment?

**Answer:**
According to the test set prediction metrics below, the Elastic Net Model has the lowest RMSE and highest R2. I would recommend to replace the original PLS model by Elastic Net model.

```{r Q6.2 model comparison}
rbind(pls_pred_metrics,
      ridge_pred_metrics,
      lasso_pred_metrics,
      enet_pred_metrics) %>%
  data.frame() %>%
  arrange(desc(Rsquared))

```


# Question 6.3
**A chemical manufacturing process for a pharmaceutical product was discussed in Sect. 1.4. In this problem, the objective is to understand the relationship between biological measurements of the raw materials (predictors), 6.5 Computing 139 measurements of the manufacturing process (predictors), and the response of product yield. Biological predictors cannot be changed but can be used to assess the quality of the raw material before processing. On the other hand, manufacturing process predictors can be changed in the manufacturing process. Improving product yield by 1% will boostrevenue by approximately one hundred thousand dollars per batch:**

## (a)
Start R and use these commands to load the data:
```{r Q6.3 load data}
data(ChemicalManufacturingProcess)

chem_predictors <- ChemicalManufacturingProcess %>% select(-Yield)  %>% as.matrix()
chem_response <- ChemicalManufacturingProcess %>% select(Yield) %>% as.matrix()
```



The matrix `processPredictors` contains the 57 predictors (12 describing the input biological material and 45 describing the process predictors) for the 176 manufacturing runs. `yield` contains the percent yield for each run.

## (b)
A small percentage of cells in the predictor set contain missing values. Use an imputation function to fill in these missing values (e.g., see Sect. 3.8).

**Anwer:**
Imputd mssing values using `missFroest` package.

```{r Q6.3 missing data imputation, warning=FALSE}
set.seed(0)
imp_chem_predictors <- missForest(chem_predictors)$ximp
```



## (c) 
Split the data into a training and a test set, pre-process the data, and tune a model of your choice from this chapter. What is the optimal value of the performance metric?


### train_test_split
```{r Q6.3 Train Test Split}
set.seed(0)

train_index <- createDataPartition(chem_response,
                                   p = 0.75,
                                   list = FALSE,
                                   times = 1) %>%
  as.vector()

data_train_X <- imp_chem_predictors[train_index,] 
data_train_Y <- chem_response[train_index,]
data_test_X <- imp_chem_predictors[-train_index,] 
data_test_Y <-chem_response[-train_index,]

```




### Build Elastic Net Model

**Answer:**
Elastic Net model is selected. The optimal fraction = 0.2928571 and lambda = 0.8571429
```{r Q6.3 Elastic Net, warning=FALSE}
set.seed(0)
enetGrid <- data.frame(.lambda = seq(0, 3, length = 50),
                       .fraction = seq(0.01, 1, length = 50))
enetTune <- train(data_train_X, data_train_Y,
                   method = 'enet',
                   tuneGrid = enetGrid,
                   trControl = ctrl,
                   preProc = c('center', 'scale'))

enetTune

```


## (d)
Predict the response for the test set.What is the value of the performance metric and how does this compare with the resampled performance metric on the training set?

**Answer:**
The R2 for the training set and test set are 0.6384729 and 0.4907965 respectively. The model has better performance on the training set.
```{r Q6.3 d}
enet_train_pred <- predict(enetTune, data_train_X)
train_metrics <- postResample(enet_train_pred,data_train_Y)
enet_test_pred <- predict(enetTune, data_test_X)
test_metrics <- postResample(enet_test_pred,data_test_Y)
rbind(train_metrics,test_metrics) %>%
  data.frame()
```

## (e)
Which predictors are most important in the model you have trained? Do either the biological or process predictors dominate the list?

**Answer:**

The top 20 important predictors are as below. The process predictors domiate the list.
```{r}
enet_vapImp <- varImp(enetTune)
enet_vapImp$importance %>% 
  arrange(desc(Overall)) %>%
  top_n(20, Overall)


```




## (f)
Explore the relationships between each of the top predictors and the response. How could this information be helpful in improving yield in future runs of the manufacturing process?

**Answer:**
As observed from the correlation plot, all biological material (BM) predictors have postive correlationship with the response variable Yield, while the manufacturing process (MP) predictors are overallly have overall smaller positive correlation with Yield than those of BMs, or have negative correslation with Yield. In future runs of manufacturing process, those individual MP predictors with small absolute value of correlation values can be further analysed and improvement actions can be taken to such MP steps in order to increase the yield so as to boost revenue.

```{r}
top_pred <- enet_vapImp$importance %>% 
  arrange(desc(Overall)) %>%
  top_n(20, Overall) %>%
  rownames_to_column() %>%
  spread(key = rowname, value = Overall)


cor_df <- data.frame(imp_chem_predictors) %>%
  select(names(top_pred)) %>%
  cbind(chem_response) 


names(cor_df) <- names(cor_df) %>%
  str_replace('BiologicalMaterial', 'BM') %>%
  str_replace('ManufacturingProcess', 'MP')

cor_df <- cor(cor_df)

corrplot(cor_df)

```
