---
title: "DATA 624 - HOMEWORK 3"
author: "FAN XU"
date: "2/28/2021"
output:
  rmdformats::readthedown:
    code_folding: hide
    df_print: paged
    highlight: tango
    number_sections: yes
    smooth_scroll: yes
    theme: united
    toc_collapsed: yes
    toc_depth: 5
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
  pdf_document:
    extra_dependencies:
    - geometry
    - multicol
    - multirow
  word_document:
    toc: yes
    toc_depth: '5'
theme: lumen
number_sections: yes
toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load library, message=FALSE, warning=FALSE}
library(tidyverse)
library(fpp2)
library(readxl)
library(rio)
library(gridExtra)
library(ggpubr)
library(ggthemes)
library(seasonal)

```



# Question - 6.2

**The plastics data set consists of the monthly sales (in thousands) of product A for a plastics manufacturer for five years.**

**(a.) Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend-cycle?**

**Answer:** 

The autoplot shows a significantinscreasing trend.
```{r 6.2 a 1}
autoplot(plastics) +
  xlab("Year") + ylab("Sales") +
  ggtitle("Autoplot: Monthly Sales of Plastic Product A") +
  theme_hc()

```

The ggsubseriesplot presents a significant normal seasonal fluctuations.

```{r 6.2 a 2}
ggsubseriesplot(plastics) +
  xlab("Year") + ylab("Sales") +
  ggtitle("ggsubseriesplot: Monthly Sales of Plastic Product A") +
  theme_hc()


```

**(b.) Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.**
```{r 6.2 b}
plastics %>% decompose(type="multiplicative") %>% 
  autoplot() + xlab("Year") +
  ggtitle("Classical Multiplicative Decomposition of Monthly Sales of Plastic Product A") +
  theme_hc()


```


**(c.) Do the results support the graphical interpretation from part a?**

**Answer:** Yes. The result of classical multiplicative decomposition supports the graphical interpretation from part A.




**(d.) Compute and plot the seasonally adjusted data.**
```{r 6.2 d, message=FALSE, warning=FALSE}
fit <- plastics %>% decompose(type="multiplicative")
autoplot(plastics, series="Original Data") +
  autolayer(trendcycle(fit), series="Trend") +
  autolayer(seasadj(fit), series="Seasonally Adjusted Data") +
  xlab("Year") + ylab("Sales") +
  ggtitle("Monthly Sales of Plastic Product A (Seasonally Adjusted)") +
  scale_color_manual(values=c("gray", "blue", "red"), 
                     breaks=c("Original Data", "Seasonally Adjusted Data", "Trend")) +
  theme_hc()


```



(e.) Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

**Anwer:**

From the autoplot after multiplicative decomposition, it is obversed that both the trend and seasonal effect got affected after introducting an outline in middle of the TS.

```{r 6.2 e 1, message=FALSE, warning=FALSE}
plastics_ol1 <- plastics
plastics_ol1[31] <- 500

fit_ol1 <- plastics_ol1 %>% decompose(type="multiplicative")
grid.arrange(
  plastics_ol1 %>% decompose(type="multiplicative") %>%
    autoplot() + xlab("Year") +
  ggtitle("Classical Multiplicative Decomposition \n(Outliner in Middle)") +
    theme_hc(),
  plastics %>% decompose(type="multiplicative") %>%
    autoplot() + xlab("Year") +
  ggtitle("Classical Multiplicative Decomposition") +
    theme_hc(),
  ncol = 2)

```

However after seasonality adjustment, the adjusted data still greatly affected by the outliner in the middle of the TS.

```{r 6.2 e 2, message=FALSE, warning=FALSE}
grid.arrange(
  autoplot(plastics_ol1, series="Original Data") +
    autolayer(trendcycle(fit_ol1), series="Trend") +
    autolayer(seasadj(fit_ol1), series="Seasonally Adjusted Data") +
    xlab("Year") + ylab("Sales") +
    ggtitle("Monthly Sales(Outliner in Middle)") +
    scale_color_manual(values=c("gray", "blue", "red"), 
                     breaks=c("Original Data", "Seasonally Adjusted Data", "Trend")) +
    theme_hc()+
    ylim(500, 1750),
  autoplot(plastics, series="Original Data") +
    autolayer(trendcycle(fit), series="Trend") +
    autolayer(seasadj(fit), series="Seasonally Adjusted Data") +
    xlab("Year") + ylab("Sales") +
    ggtitle("Monthly Sales") +
    scale_color_manual(values=c("gray", "blue", "red"), 
                     breaks=c("Original Data", "Seasonally Adjusted Data", "Trend")) +
    theme_hc() +
    ylim(500, 1750),
  ncol = 2)

```


**(f.) Does it make any difference if the outlier is near the end rather than in the middle of the time series?**

**Answer:** The outlier near the end has less effect on both the trend and seasonal fluctuation than that in the middle of the TS.

```{r 6.2 f 1, message=FALSE, warning=FALSE}
plastics_ol2 <- plastics
plastics_ol2[58] <- 500

fit_ol2 <- plastics_ol2 %>% decompose(type="multiplicative")
grid.arrange(
  plastics_ol2 %>% decompose(type="multiplicative") %>%
    autoplot() + xlab("Year") +
  ggtitle("Classical Multiplicative Decomposition \n(Outliner at the end)") +
    theme_hc(),
  plastics %>% decompose(type="multiplicative") %>%
    autoplot() + xlab("Year") +
  ggtitle("Classical Multiplicative Decomposition") +
    theme_hc(),
  ncol = 2)

```





```{r 6.2 f 2, message=FALSE, warning=FALSE}
grid.arrange(
  autoplot(plastics_ol2, series="Original Data") +
    autolayer(trendcycle(fit_ol2), series="Trend") +
    autolayer(seasadj(fit_ol2), series="Seasonally Adjusted Data") +
    xlab("Year") + ylab("Sales") +
    ggtitle("Monthly Sales(Outliner in Middle)") +
    scale_color_manual(values=c("gray", "blue", "red"), 
                     breaks=c("Original Data", "Seasonally Adjusted Data", "Trend")) +
    theme_hc()+
    ylim(500, 1750),
  autoplot(plastics, series="Original Data") +
    autolayer(trendcycle(fit), series="Trend") +
    autolayer(seasadj(fit), series="Seasonally Adjusted Data") +
    xlab("Year") + ylab("Sales") +
    ggtitle("Monthly Sales") +
    scale_color_manual(values=c("gray", "blue", "red"), 
                     breaks=c("Original Data", "Seasonally Adjusted Data", "Trend")) +
    theme_hc() +
    ylim(500, 1750),
  ncol = 2)

```


# Question - 6.3

**Recall your retail time series data (from Exercise 3 in Section 2.10).**

**Decompose the series using X11. Does it reveal any outliers, or unusual features that you had not noticed previously?**


**Answer:**

From the X11 decomposition remainder graph, it is observed there are multiple up or down spikes, which are possible outliners.



```{r 6.3 1}
data <- import("https://raw.githubusercontent.com/shirley-wong/Data-624/main/HW1/retail.xlsx", skip=1) 


myts <- ts(data[,"A3349398A"], frequency=12, start=c(1982,4))
autoplot(myts) +
  ggtitle("Monthly Food Retailing in Australia") +
  xlab("Year") + 
  ylab("Sales") +
  theme_hc()

```


```{R 6.3 2}
#x11 decomposition
fit <- myts %>% seas(x11="")
autoplot(fit, series="Data") +
  xlab("Year") + ylab("Sales") +
  ggtitle("X11 Decomposition of Monthly Food Retailing in Australia")

```
